Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-22 paper/intro.tex
Minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@673 3fc257bb-8550-0410-81b3-a5c103da83f7

More recently, MapReduce formulations of traditional number-crunching
kinds of scientific computational tasks have been described, such as
post-processing analysis of simulation data~\cite{Tu08}, graph
algorithmics~\cite{Cohen09}, and linear algebra
operations~\cite{Ekanayake09}.  The paper by Tu et al.~\cite{Tu08} was
particularly insightful to us, because it described how MapReduce
could be implemented on top of the ubiquitous distributed-memory
message-passing interface (MPI), and how the intermediate data-shuffle
operation is conceptually identical to the familiar MPI\_Alltoall
operation.  Their implementation of MapReduce was within a Python
wrapper to simplify the writing of user programs.  The paper motivated
us to develop our own C++ library built on top of MPI for use in graph
analytics, which we initially released as open-source software in
mid-2009.  We have since worked to optimize several of the library's
underlying algorithms and to enable its operation in out-of-core mode
for larger data sets.  These algorithmic improvements are described in
this paper and are part of the current downloadable version; see
Section ~\ref{sec:thanks} for details.

@@

\item {\it No fault tolerance or data redundancy:} Current MPI
implementations do not enable easy detection of a dead processor or
retrieval of the data it was working on.  So like most MPI programs, a
parallel program calling the MR-MPI library will hang or crash if a
processor goes away.  Unlike Hadoop, and its file system (HDFS) which
provides data redundancy, the MR-MPI library reads and writes simple,
flat files.  It can use local per-processor disks, or a parallel file
system, if available, but these typically provide no data redundancy.

@@

The remainder of the paper is organized as follows.
Sections~\ref{sec:mr} and~\ref{sec:outcore} describe how in-core and
out-of-core MapReduce primitives are formulated as MPI-based
operations in the MR-MPI library.  Section~\ref{sec:graph} briefly
describes the formulation of several common graph algorithms as
MapReduce operations.  Section~\ref{sec:results} gives performance
results for these algorithms running on a parallel cluster for graphs
ranging in size from 8 million to 2 billion edges; we highlight the
performance and complexity trade-offs of a MapReduce approach versus
other more special-purpose algorithms.  The latter generally perform
better but are harder to implement efficiently on distributed memory
machines, due to the required explicit management of parallelism,
particularly for large out-of-core data sets.
Section~\ref{sec:lessons} summarizes some lessons learned from the
implementation and use of our library.

>>>>

More recently, MapReduce formulations of traditional number-crunching
kinds of scientific computational tasks have been described, such as
post-processing analysis of simulation data~\cite{Tu08}, graph
algorithmics~\cite{Cohen09}, and linear algebra
operations~\cite{Ekanayake09}.  The paper by Tu et al.~\cite{Tu08} was
particularly insightful to us, because it described how MapReduce
could be implemented on top of the ubiquitous distributed-memory
message-passing interface (MPI), and how the intermediate data-shuffle
operation is conceptually identical to the familiar MPI\_Alltoall
operation.  Their implementation of MapReduce was within a Python
wrapper to simplify the writing of user programs.  The paper motivated
us to develop our own C++ library built on top of MPI for use in graph
analytics, which we initially released as open-source software in
mid-2009.  We have since worked to optimize several of the library's
underlying algorithms and to enable its operation in out-of-core mode
for larger data sets.  These algorithmic improvements are described in
this paper and are part of the current downloadable version; see
Section~\ref{sec:thanks} for details.

@@

\item {\it No fault tolerance or data redundancy:} Current MPI
implementations do not enable easy detection of a dead processor or
retrieval of the data it was working on.  So like most MPI programs, a
parallel program calling the MR-MPI library will hang or crash if a
processor goes away.  Unlike Hadoop, and its file system (HDFS) which
provides data redundancy, the MR-MPI library reads and writes simple,
flat files.  It can use local per-processor disks, or a parallel file
system, if available, but these typically do not provide data redundancy.

@@

The remainder of the paper is organized as follows.
Sections~\ref{sec:mr} and~\ref{sec:outcore} describe how in-core and
out-of-core MapReduce primitives are formulated as MPI-based
operations in the MR-MPI library.  Section~\ref{sec:graph} briefly
describes the formulation of several common graph algorithms as
MapReduce operations.  Section~\ref{sec:results} gives performance
results for these algorithms running on a parallel cluster for graphs
ranging in size from 8 million to 2 billion edges. We highlight the
performance and complexity trade-offs of a MapReduce approach versus
other more special-purpose algorithms.  The latter generally perform
better but are harder to implement efficiently on distributed memory
machines, due to the required explicit management of parallelism,
particularly for large out-of-core data sets.
Section~\ref{sec:lessons} summarizes some lessons learned from the
implementation and use of our library.

### 2010-11-22 paper/mr.tex
Minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@673 3fc257bb-8550-0410-81b3-a5c103da83f7

The {\it collate()} operation (or data shuffle in Hadoop) identifies
unique keys and collects all the values associated with those keys to
create KMV pairs.  This is done in two stages, the first of which
requires communication, since KV pairs with the same key may be owned
by any processor.  Each processor hashes each of its keys to determine
which processor will ``own'' it.  The $k$-byte length key is hashed
into a 32-bit value whose remainder modulo $P$ processors is the
owning processor rank.  Alternatively, the user can provide a hash
function which converts the key into a processor rank.  Each processor
then sends each of its KV pairs to the owning processor.

@@

The interface to various low-level operations is provided so that a
user's program can string them together in various ways to construct
useful MapReduce algorithms.  For example, output from a {\it
reduce()} can serve as input to a subsequent {\it map()} or {\it
collate()}.  Likewise, data can be partitioned across multiple
MapReduce (MR) objects, each with their own set of KV pairs.  For
example, as will be illustrated in Section ~\ref{sec:graph}, one MR
object can store the edges of a graph, and another its vertices.  KV
pairs from multiple MR objects can be combined to perform new
sequences of {\it map()}, {\it collate()}, and {\it reduce()}
operations.

>>>>

The {\it collate()} operation (or data shuffle in Hadoop) identifies
unique keys and collects all the values associated with those keys to
create KMV pairs.  This is done in two stages, the first of which
requires communication, since KV pairs with the same key may be owned
by any processor.  Each processor hashes each of its keys to determine
which processor will ``own'' it.  The $k$-byte length key is hashed
into a 32-bit value whose remainder modulo $P$ processors is the
owning processor rank.  Alternatively, the user can provide a hash
function that converts the key into a processor rank.  Each processor
then sends each of its KV pairs to the owning processor.

@@

The interface to various low-level operations is provided so that a
user's program can string them together in various ways to construct
useful MapReduce algorithms.  For example, output from a {\it
reduce()} can serve as input to a subsequent {\it map()} or {\it
collate()}.  Likewise, data can be partitioned across multiple
MapReduce (MR) objects, each with its own set of KV pairs.  For
example, as will be illustrated in Section~\ref{sec:graph}, one MR
object can store the edges of a graph, and another its vertices.  KV
pairs from multiple MR objects can be combined to perform new
sequences of {\it map()}, {\it collate()}, and {\it reduce()}
operations.

### 2010-11-22 paper/outcore.tex
Minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@673 3fc257bb-8550-0410-81b3-a5c103da83f7

Performing an our-of-core {\it collate()} operation is more complex.
Recall that the operation occurs in two stages.  First, keys are
hashed to ``owning'' processors and the KV pairs are communicated to
new processors in an all-to-all fashion.  In out-of-core mode, this
operation is performed on one page of KV data (per processor) at a
time.  Each processor reads in a page of KV pairs, the communication
pattern is determined (which processors receive which datums in that
page), and all-to-all communication is performed.  Each processor
allocates a two-page chunk of memory to receive incoming KV pairs.  On
average each processor should receive one page of KV pairs; the
two-page allocation allows for some load imbalance.  If the KV pairs
are distributed unevenly so that this limit is exceeded, the
all-to-all communication is performed in smaller, multiple passes
until the full page contributed by every processor has been
communicated.  Performing an MPI\_Alltoall(), or using the custom
all-to-all routines provided by the MR-MPI library, requires
allocation of auxiliary arrays that store processor indices, datum
pointers, and datum lengths.  For the case of many tiny KV pairs, the
total number of memory pages required to perform the all-to-all
communication, including the source and target KV pages is at most
seven.

@@

\begin{figure}
\includegraphics[width=\textwidth,angle=-90]{fig_collate2.pdf}
\caption{\it Multi-pass algorithm for converting KV data to KMV data.
The vertical lines represent out-of-core data sets.  KV is key/value
pairs; HT is an in-memory hash table; KMV is key/multivalue pairs.}
\label{fig:collate}
\end{figure}

>>>>

Performing an out-of-core {\it collate()} operation is more complex.
Recall that the operation occurs in two stages.  First, keys are
hashed to ``owning'' processors and the KV pairs are communicated to
new processors in an all-to-all fashion.  In out-of-core mode, this
operation is performed on one page of KV data (per processor) at a
time.  Each processor reads in a page of KV pairs, the communication
pattern is determined (which processors receive which datums in that
page), and all-to-all communication is performed.  Each processor
allocates a two-page chunk of memory to receive incoming KV pairs.  On
average each processor should receive one page of KV pairs; the
two-page allocation allows for some load imbalance.  If the KV pairs
are distributed unevenly so that this limit is exceeded, the
all-to-all communication is performed in multiple smaller passes
until the full page contributed by every processor has been
communicated.  Performing an MPI\_Alltoall(), or using the custom
all-to-all routines provided by the MR-MPI library, requires
allocation of auxiliary arrays that store processor indices, datum
pointers, and datum lengths.  For the case of many tiny KV pairs, the
total number of memory pages required to perform the all-to-all
communication, including the source and target KV pages is at most
seven.

@@

\begin{figure}
\includegraphics[width=\textwidth,angle=-90]{fig_collate2.pdf}
\caption{\it Multi-pass algorithm for converting KV data to KMV data within
a processor.
The vertical lines represent out-of-core data sets.  KV is key/value
pairs; HT is an in-memory hash table; KMV is key/multivalue pairs.}
\label{fig:collate}
\end{figure}

### 2010-11-22 paper/paper.tex
Minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@673 3fc257bb-8550-0410-81b3-a5c103da83f7

We describe a parallel library written with message-passing (MPI)
calls that allows algorithms to be expressed in the MapReduce
paradigm.  This means the calling program does not need to include
explicit parallel code, but instead provides ``map'' and ``reduce''
functions that operate independently on elements of a data set
distributed across processors.  The library performs needed data
movement between processors.  We describe how typical MapReduce
functionality can be implemented in an MPI context, and also in an
out-of-core manner for data sets that do not fit within the aggregate
memory of a parallel machine.  Our motivation for creating this
library was to enable graph algorithms to be written as MapReduce
operations, allowing processing of terabyte-scale data sets.  We
outline MapReduce versions of several such algorithms: vertex ranking
via PageRank, triangle finding, connected component identification,
Luby's algorithm for maximally independent sets, and single-source
shortest-path calculation.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized R-MAT matrices in
parallel; a MapReduce version of this operation is also described.
Performance and scalability results for the various algorithms are
presented for varying size graphs on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.

>>>>

We describe a parallel library written with message-passing (MPI)
calls that allows algorithms to be expressed in the MapReduce
paradigm.  This means the calling program does not need to include
explicit parallel code, but instead provides ``map'' and ``reduce''
functions that operate independently on elements of a data set
distributed across processors.  The library performs needed data
movement between processors.  We describe how typical MapReduce
functionality can be implemented in an MPI context, and also in an
out-of-core manner for data sets that do not fit within the aggregate
memory of a parallel machine.  Our motivation for creating this
library was to enable graph algorithms to be written as MapReduce
operations, allowing processing of terabyte-scale data sets on 
traditional MPI-based clusters.  We
outline MapReduce versions of several such algorithms: vertex ranking
via PageRank, triangle finding, connected component identification,
Luby's algorithm for maximally independent sets, and single-source
shortest-path calculation.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized R-MAT matrices in
parallel; a MapReduce version of this operation is also described.
Performance and scalability results for the various algorithms are
presented for varying size graphs on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.
